Ring of Fire Camera Detections Report

### 2023-03-02 analysis/paper/paper.qmd
updates

cams_sf <- RoF_CameraDeploymentLog_WLU %>% 
  dplyr::filter(!is.na(latitude)) %>% 
  dplyr::filter(!is.na(longitude)) %>% 
  sf::st_as_sf(coords=c("longitude","latitude"), crs="EPSG:4326") %>% 
  dplyr::mutate(is_retrieved = !is.na(date_retrieved))

# can <- rnaturalearth::ne_countries(country = "Canada", returnclass = "sf") %>% 
#   sf::st_crop(cams_sf)
# 
# can %>% 
#   ggplot2::ggplot() +
#   ggplot2::theme_bw() +
#   ggplot2::geom_sf() +
#   ggplot2::geom_sf(data=cams_sf)

cams_sf %>%
  ggplot2::ggplot() +
  ggplot2::theme_bw() +
  ggplot2::geom_sf(ggplot2::aes(color = is_retrieved))
```

@@

# Join
anns_wide_animals_joined <- anns_wide_animals %>% 
  dplyr::right_join(dets_images, by = "image_id")
```

@@

extr_16 <- exactextractr::exact_extract(lu_16, cams_sf_buffers_16)
extr_17 <- exactextractr::exact_extract(lu_17, cams_sf_buffers_17)

@@

```{r}
#| label: ordination-base
#| echo: false

@@

test_rda <- vegan::rda(anns_wide_animals_table_tb ~ .,
                       extr_table_sub_tb)

plot(test_rda,
     display = c("sp", "lc", "cn")
)
```

```{r}
#| label: ordination-custom
#| echo: false

custom_rda_plot(test_rda)
```

@@

events_tb <- anns_wide_animals_joined %>% 
  dplyr::filter(species != "Unknown") %>%
  dplyr::mutate(plot_id = stringr::str_sub(id, 1, 4)) %>% 
  dplyr::mutate(date_time = lubridate::ymd_hms(exif_date_time_original,
                                               tz = "Canada/Eastern"),
                year = lubridate::year(date_time),
                month = lubridate::month(date_time),
                day = lubridate::day(date_time),
                ymd = paste(year, month, day, sep = "_")) %>% 
  # dplyr::mutate(ind = paste(species, sex, age, sep = "_")) %>% 
  # dplyr::select(camera_id, image_id, species, ymd) %>% 
  dplyr::select(plot_id, species, ymd) %>% 
  # dplyr::group_by(image_id, ind, ymd) %>% 
  # dplyr::summarise(ind_count=dplyr::n())
  dplyr::distinct() %>% 
  dplyr::group_by(plot_id, species) %>%
  dplyr::summarise(day_count=dplyr::n()) %>% 
  dplyr::ungroup() %>% 
  # dplyr::group_by(camera_id) %>% 
  # dplyr::mutate(day_prop = day_count/sum(day_count)) # %>% 
  tidyr::pivot_wider(values_from = "day_count", names_from = "species") %>%
  dplyr::arrange(plot_id) %>%
  tibble::column_to_rownames("plot_id")

events_tb[is.na(events_tb)] <- 0
events_tb <- events_tb[,colSums(events_tb)>50]
tmp_col <- ncol(events_tb)

# events <- events/rowSums(events)
events <- events_tb %>% 
  tibble::rownames_to_column("plot_id")

```

@@

```{r}
#| label: ordination-events
#| echo: false

anns_wide_animals_joined_mut <- anns_wide_animals_joined %>% 
  dplyr::filter(species != "Unknown") %>%
  dplyr::mutate(plot_id = stringr::str_sub(id, 1, 4)) %>% 
  dplyr::mutate(date_time = lubridate::ymd_hms(exif_date_time_original,
                                               tz = "Canada/Eastern"),
                year = lubridate::year(date_time),
                month = lubridate::month(date_time),
                day = lubridate::day(date_time),
                ymd = paste(year, month, day, sep = "_"))

events_tb_cam <- anns_wide_animals_joined_mut %>% 
  # dplyr::mutate(ind = paste(species, sex, age, sep = "_")) %>% 
  # dplyr::select(camera_id, image_id, species, ymd) %>% 
  dplyr::select(camera_id, species, ymd) %>% 
  # dplyr::group_by(image_id, ind, ymd) %>% 
  # dplyr::summarise(ind_count=dplyr::n())
  dplyr::distinct() %>% 
  dplyr::group_by(camera_id, species) %>%
  dplyr::summarise(day_count=dplyr::n()) %>% 
  dplyr::ungroup() %>% 
  # dplyr::group_by(camera_id) %>% 
  # dplyr::mutate(day_prop = day_count/sum(day_count)) # %>% 
  tidyr::pivot_wider(values_from = "day_count", names_from = "species") %>%
  dplyr::arrange(camera_id) %>%
  tibble::column_to_rownames("camera_id")

events_tb_cam[is.na(events_tb_cam)] <- 0
# events_tb_cam <- events_tb_cam[,colSums(events_tb_cam)>50]
# tmp_col <- ncol(events_tb_cam)

@@

events_rda <- vegan::rda( events_tb_cam_hell ~ .,
                          extr_table_sub_tb)

# vegan::RsquareAdj(events_rda)

custom_rda_plot(events_rda, sp_scale = 1.5)
```

@@

events_gower <- cluster::daisy(events_tb_cam_hell, "gower")
clust <- hclust(events_gower, method = "ward.D2")
plot(clust)
```

@@

clust_grp <- cutree(clust, k=5)
custom_rda_plot(events_rda, sp_scale = 1.5, clust = clust_grp) 
```

@@

cams_sf_with_props <- cams_sf_coords %>% 
  sf::st_drop_geometry() %>% 
  dplyr::select(plot_id,X,Y) %>% 
  dplyr::group_by(plot_id) %>% 
  dplyr::mutate(X=mean(X), Y=mean(Y)) %>%
  dplyr::distinct() %>% 
  dplyr::ungroup() %>% 
  dplyr::left_join(events, by = "plot_id")

ggplot2::ggplot() +
  scatterpie::geom_scatterpie(data = cams_sf_with_props,
                              ggplot2::aes(x=X, y=Y, group=plot_id),
                              cols=names(cams_sf_with_props)[4:(4+tmp_col-1)]) + 
  ggplot2::coord_equal() #+ 
#ggplot2::theme(legend.position = "none")
```

@@

zones <- 
  sf::st_read("analysis/data/raw_data/ecozones/ECOREGN/LIO-2022-10-25/ECOREGION.shp") %>% 
  sf::st_transform(crs = sf::st_crs(cams_sf))

plot(zones$geometry)
```

@@

zones_filter <- zones %>%
  sf::st_filter(cams_sf)
plot(zones_filter)
```

ggplot2::ggplot(zones_filter) +
  ggplot2::theme_bw() +
  ggplot2::geom_sf()
```

@@

ggplot2::ggplot(zones_filter) +
  ggplot2::theme_bw() +
  ggplot2::geom_sf() +
  ggplot2::geom_sf(data = cams_sf, cex = 0.5, 
                   ggplot2::aes(col=is_retrieved)) +
  ggplot2::coord_sf(xlim = c(-89, -84),
                    ylim = c(50, 55))
```

@@

ggplot2::ggplot(zones_filter) +
  ggplot2::theme_bw() +
  ggplot2::geom_sf() +
  ggplot2::geom_sf(data = cams_sf_clust, cex = 1, 
                   ggplot2::aes(col = grp)) +
  ggplot2::coord_sf(xlim = c(-89, -84),
                    ylim = c(50, 55))
```

@@

clust_geo <- ClustGeo::hclustgeo(events_gower_mod,
                                 dist_geo, alpha = 1)
clust_geo_grp <- cutree(clust_geo, k=K)

@@

plot(clust_geo)
```

@@

ggplot2::ggplot(zones_filter) +
  ggplot2::theme_bw() +
  ggplot2::geom_sf() +
  ggplot2::geom_sf(data = cams_sf_geo_clust, cex = 1, 
                   ggplot2::aes(col = grp)) +
  ggplot2::coord_sf(xlim = c(-89, -84),
                    ylim = c(50, 55))
```

@@

custom_rda_plot(events_rda, sp_scale = 1.5, clust = clust_geo_grp)
```

@@

cams_sf_sp <- anns_wide_animals_joined_mut %>% 
  dplyr::left_join(cams_sf_coords, by = dplyr::join_by(camera_id, plot_id)) %>% 
  dplyr::filter(species %in% colnames(events[,-1]))

cams_sf_sp_sum <- cams_sf_sp %>% 
  dplyr::group_by(camera_id, X, Y, species, month) %>% 
  dplyr::summarise(count=dplyr::n()) %>% 
  dplyr::ungroup()
# stopifnot(dim(cams_sf_sp)[1]==dim(anns_wide_animals_joined_mut)[1])

# cams_sf_sp %>% 
cams_sf_sp_sum %>% 
  # dplyr::filter(species %in% colnames(events[,-1])) %>% 
  dplyr::filter(species == "Caribou") %>% 
  # dplyr::filter(species == "Caribou", month == 8) %>% 
  ggplot2::ggplot() +
  # ggplot2::geom_hex(ggplot2::aes(x=X, y=Y), bins = 70) +
  # ggplot2::stat_density_2d(ggplot2::aes(fill = ..level..), geom = "polygon", colour="white") +
  # ggplot2::geom_density_2d(ggplot2::aes(x=X, y=Y)) +
  # ggplot2::stat_density_2d(ggplot2::aes(x=X, y=Y, fill = count),
  #                          geom = "raster", contour = T) +
  # ggplot2::stat_density_2d(ggplot2::aes(x=X, y=Y, fill = count),
  #                          geom = "raster", contour = F) +
  # ggplot2::geom_tile(ggplot2::aes(x=X, y=Y, fill = count),interpolate=TRUE) +
  ggplot2::geom_point(ggplot2::aes(x=X, y=Y, size = count)) +
  # ggplot2::scale_fill_continuous(type = "viridis") +
  # ggplot2::scale_fill_distiller(palette=4, direction=1) +
  # ggplot2::theme_bw() + 
  ggplot2::facet_wrap(~month) #+
  # ggplot2::scale_x_continuous(expand = c(0, 0)) +
  # ggplot2::scale_y_continuous(expand = c(0, 0)) # +
  # ggplot2::theme(
  #   legend.position='none'
  # )
```

```{r}
ggplot2::ggplot(ggplot2::faithfuld, ggplot2::aes(waiting, eruptions)) +
 ggplot2::geom_raster(ggplot2::aes(fill = density))
```

@@

-->

>>>>

cams_sf <- RoF_CameraDeploymentLog_WLU %>% 
  dplyr::filter(!is.na(latitude)) %>% 
  dplyr::filter(!is.na(longitude)) %>% 
  sf::st_as_sf(coords=c("longitude","latitude"), crs="EPSG:4326") %>% 
  dplyr::mutate(is_retrieved = ifelse(!is.na(date_retrieved), "Yes", "No"))

# can <- rnaturalearth::ne_countries(country = "Canada", returnclass = "sf") %>% 
#   sf::st_crop(cams_sf)
# 
# can %>% 
#   ggplot2::ggplot() +
#   ggplot2::theme_bw() +
#   ggplot2::geom_sf() +
#   ggplot2::geom_sf(data=cams_sf)
```

@@

# Joins

anns_wide_animals_joined_mut <- anns_wide_animals_joined %>% 
  dplyr::filter(species != "Unknown") %>%
  dplyr::mutate(plot_id = stringr::str_sub(id, 1, 4)) %>% 
  dplyr::mutate(date_time = lubridate::ymd_hms(exif_date_time_original,
                                               tz = "Canada/Eastern"),
                year = lubridate::year(date_time),
                month = lubridate::month(date_time, label = TRUE),
                day = lubridate::day(date_time),
                ymd = paste(year, month, day, sep = "_"))

# anns_wide_animals_joined_dets <- anns_wide_animals %>% 
#   dplyr::right_join(dets_images, by = "image_id")
```

@@

extr_16 <- exactextractr::exact_extract(lu_16, cams_sf_buffers_16,
                                        progress = FALSE)
extr_17 <- exactextractr::exact_extract(lu_17, cams_sf_buffers_17,
                                        progress = FALSE)

@@

```{r}
#| label: RDA-presence-absence
#| echo: false

@@

pa_rda <- vegan::rda(anns_wide_animals_table_tb ~ .,
                       extr_table_sub_tb)

custom_rda_plot(pa_rda)
```

@@

events_plot_days <- anns_wide_animals_joined_mut %>% 
  # dplyr::mutate(ind = paste(species, sex, age, sep = "_")) %>% 
  # dplyr::select(camera_id, image_id, species, ymd) %>% 
  dplyr::select(plot_id, species, ymd) %>% 
  # dplyr::group_by(image_id, ind, ymd) %>% 
  # dplyr::summarise(ind_count=dplyr::n())
  dplyr::distinct()

events_plot <- events_plot_days %>% 
  dplyr::group_by(plot_id, species) %>%
  dplyr::summarise(day_count=dplyr::n()) %>% 
  dplyr::ungroup()

events_plot_tb <- events_plot %>% 
  # dplyr::group_by(camera_id) %>% 
  # dplyr::mutate(day_prop = day_count/sum(day_count)) # %>% 
  tidyr::pivot_wider(values_from = "day_count", names_from = "species") %>%
  dplyr::arrange(plot_id) %>%
  tibble::column_to_rownames("plot_id")

events_plot_tb[is.na(events_plot_tb)] <- 0
events_plot_tb_filt <- events_plot_tb[,colSums(events_plot_tb)>50]
tmp_col <- ncol(events_plot_tb_filt)

# events <- events/rowSums(events)
events_plot_tb_filt_with_col <- events_plot_tb_filt %>% 
  tibble::rownames_to_column("plot_id")
```

@@

```{r}
#| label: RDA-events
#| echo: false

events_cam_days <- anns_wide_animals_joined_mut %>% 
  # dplyr::mutate(ind = paste(species, sex, age, sep = "_")) %>% 
  # dplyr::select(camera_id, image_id, species, ymd) %>% 
  dplyr::select(camera_id, species, month, ymd) %>% 
  # dplyr::group_by(image_id, ind, ymd) %>% 
  # dplyr::summarise(ind_count=dplyr::n())
  dplyr::distinct()

events_cam <- events_cam_days %>% 
  dplyr::group_by(camera_id, species) %>%
  dplyr::summarise(day_count=dplyr::n()) %>% 
  dplyr::ungroup()

events_tb_cam <- events_cam %>% 
  # dplyr::group_by(camera_id) %>% 
  # dplyr::mutate(day_prop = day_count/sum(day_count)) # %>% 
  tidyr::pivot_wider(values_from = "day_count", names_from = "species") %>%
  dplyr::arrange(camera_id) %>%
  tibble::column_to_rownames("camera_id")

events_tb_cam[is.na(events_tb_cam)] <- 0
# events_tb_cam_filt <- events_tb_cam[,colSums(events_tb_cam)>50]
# tmp_col <- ncol(events_tb_cam)

@@

events_rda <- vegan::rda(events_tb_cam_hell ~ .,
                         extr_table_sub_tb)

# vegan::RsquareAdj(events_rda)
# custom_rda_plot(events_rda, sp_scale = 1.5, scale_x_sp=0.4, scale_y_sp=0.3,
#                 thres_plot = 0.3)
```

@@

events_gower <- cluster::daisy(events_tb_cam_hell, "gower")
clust <- hclust(events_gower, method = "ward.D2")
# plot(clust)
```

@@

clust_grp <- cutree(clust, k=5)
custom_rda_plot(events_rda, sp_scale = 1.5, scale_x_sp=0.4, scale_y_sp=0.3,
                thres_plot = 0.3, clust = clust_grp)
```

@@

cams_sf_with_props <- cams_sf_coords %>% 
  sf::st_drop_geometry() %>% 
  dplyr::select(plot_id,X,Y) %>% 
  dplyr::group_by(plot_id) %>% 
  dplyr::mutate(X=mean(X), Y=mean(Y)) %>%
  dplyr::distinct() %>% 
  dplyr::ungroup() %>% 
  dplyr::left_join(events_plot_tb_filt_with_col, by = "plot_id")

# ggplot2::ggplot() +
#   scatterpie::geom_scatterpie(data = cams_sf_with_props,
#                               ggplot2::aes(x=X, y=Y, group=plot_id),
#                               cols=names(cams_sf_with_props)[4:(4+tmp_col-1)]) + 
#   ggplot2::coord_equal() #+ 
#ggplot2::theme(legend.position = "none")
```

@@

zones <- 
  sf::st_read("analysis/data/raw_data/ecozones/ECOREGN/LIO-2022-10-25/ECOREGION.shp",
              quiet=TRUE) %>% 
  sf::st_transform(crs = sf::st_crs(cams_sf))

# plot(zones$geometry)
```

@@

zones_filter <- zones %>%
  sf::st_filter(cams_sf)
# plot(zones_filter)
```

@@

ggplot2::ggplot() +
  ggplot2::theme_bw() +
  ggplot2::geom_sf(data = zones_filter) +
  ggplot2::geom_sf(data = sf::st_jitter(cams_sf, 0.05), cex = 0.5, 
                   ggplot2::aes(col=is_retrieved)) +
  ggplot2::coord_sf(xlim = c(-89, -84),
                    ylim = c(50, 55)) +
  ggplot2::labs(col = "Camera \nRetrieved")
```

@@

ggplot2::ggplot(zones_filter) +
  ggplot2::theme_bw() +
  ggplot2::geom_sf() +
  ggplot2::geom_sf(data = sf::st_jitter(cams_sf_clust, 0.05), cex = 1, 
                   ggplot2::aes(col = grp)) +
  ggplot2::coord_sf(xlim = c(-89, -84),
                    ylim = c(50, 55))
```

@@

clust_geo <- ClustGeo::hclustgeo(events_gower_mod,
                                 dist_geo, alpha = 0.2)
clust_geo_grp <- cutree(clust_geo, k=K)

@@

# plot(clust_geo)
```

@@

ggplot2::ggplot(zones_filter) +
  ggplot2::theme_bw() +
  ggplot2::geom_sf() +
  ggplot2::geom_sf(data = sf::st_jitter(cams_sf_geo_clust, 0.05), cex = 1, 
                   ggplot2::aes(col = grp)) +
  ggplot2::coord_sf(xlim = c(-89, -84),
                    ylim = c(50, 55))
```

@@

# custom_rda_plot(events_rda, sp_scale = 1.5, clust = clust_geo_grp)
```

@@

# cams_sf_sp <- anns_wide_animals_joined_mut %>% 
#   dplyr::left_join(cams_sf_coords, by = dplyr::join_by(camera_id, plot_id)) %>% 
#   dplyr::filter(species %in% colnames(events[,-1]))

cams_sf_sp <- events_cam_days %>% 
  dplyr::left_join(cams_sf_coords, by = dplyr::join_by(camera_id)) %>% 
  dplyr::filter(species %in% colnames(events_plot_tb_filt_with_col[,-1]))

cams_sf_sp_sum <- cams_sf_sp %>% 
  dplyr::group_by(camera_id, X, Y, geometry, species, month) %>% 
  dplyr::summarise(count=dplyr::n()) %>% 
  dplyr::ungroup() # %>% 
  # dplyr::group_by(camera_id, X, Y, species) %>% 
  # dplyr::mutate(count=count/sum(count)) %>% 
  # dplyr::ungroup() 
# stopifnot(dim(cams_sf_sp)[1]==dim(anns_wide_animals_joined_mut)[1])
```

```{r}
cams_sf_sp %>%
  dplyr::filter(species == "Caribou", month == "Jun") %>%
  ggplot2::ggplot(ggplot2::aes(X, Y)) +
  ggplot2::geom_density_2d_filled(binwidth = 0.05) +
  ggplot2::coord_equal(xlim = c(-89, -84),
                       ylim = c(50, 55)) +
  ggplot2::theme(
    legend.position='none'
  )
```

```{r}
cams_sf_sp %>%
  dplyr::filter(species == "Caribou") %>%
  ggplot2::ggplot(ggplot2::aes(X, Y)) +
  ggplot2::geom_density_2d_filled(breaks = c(seq(0,1,by=0.1),5, 10, 20, 50,100),
                                  n=20) +
  ggplot2::facet_wrap(~month) +
  ggplot2::coord_equal(xlim = c(-89, -84),
                       ylim = c(50, 55)) +
  ggplot2::theme(
    legend.position='none'
  )
```

```{r}
# cams_sf_sp_sum %>%
#   # dplyr::filter(species == "Caribou") %>%
#   ggplot2::ggplot() +
#   ggplot2::geom_point(ggplot2::aes(x=X, y=Y, size=count, col =species)) +
#   ggplot2::facet_wrap(~month) +
#   ggplot2::coord_equal(xlim = c(-89, -84),
#                        ylim = c(50, 55))

for (sp in unique(cams_sf_sp_sum$species)){

  dat <- sf::st_as_sf(cams_sf_sp_sum) %>% 
    dplyr::filter(species == sp)

  gg <- ggplot2::ggplot(zones_filter) +
    ggplot2::theme_bw() +
    ggplot2::geom_sf() +
    ggplot2::geom_sf(data = dat,
                     ggplot2::aes(size = count)) +
    ggplot2::coord_sf(xlim = c(-89, -84),
                      ylim = c(50, 54)) +
    ggplot2::facet_wrap(~month) +
    ggplot2::labs(size="Days \ndetected")

  ggplot2::ggsave(gg, filename = paste0("analysis/figures/",sp,".png"))
}
```

```{r}
# cams_sf_sp_sum %>%
#   dplyr::filter(species == "Caribou") %>%
#   ggplot2::ggplot() +
#   ggplot2::geom_raster(ggplot2::aes(x=X, y=Y, z=count)) +
#   ggplot2::facet_wrap(~month) +
#   ggplot2::coord_equal(xlim = c(-89, -84),
#                        ylim = c(50, 55))
```

@@

# ggplot2::ggplot(zones_filter) +
#   ggplot2::theme_bw() +
#   ggplot2::geom_sf()
```
-->
